\documentclass{beamer}
\usetheme{Madrid}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\title[Semantic Search Challenges]{"Challenges in Semantic Search: User Intent and Context"}
\author{Aliaksei Zimnitski, MSc. Mirwais Ahmadzai}
\institute[STU]{Slovenská technická univerzita v Bratislave}
\date{\today}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Abstract}
Semantic search is a vital technology in the field of information retrieval, playing a pivotal role in enhancing search accuracy by focusing on understanding user intent and context. This presentation delves into the challenges faced by semantic search systems.
\end{frame}

\begin{frame}
\frametitle{Introduction}
Semantic search revolutionizes information retrieval by focusing on the meaning behind queries. This approach goes beyond keyword matching, emphasizing user intent and context.
\newline
\newline
This comparison highlights key differences in their capabilities.
\begin{table}
\centering
\small % Makes the font smaller
\setlength{\tabcolsep}{3pt} % Reduces the space between columns
\renewcommand{\arraystretch}{1.2} % Adjusts the space between rows
\begin{tabular}{|l|l|l|}
\hline
\textbf{Feature} & \textbf{Trad. Search} & \textbf{Sem. Search} \\ \hline
Keyword Dependency & High & Low \\ \hline
Context Understanding & Low & High \\ \hline
User Intent Interpretation & Limited & Advanced \\ \hline
Accuracy in Ambiguous Queries & Low & High \\ \hline
Adaptability to Language Evolution & Low & High \\ \hline
\end{tabular}
\caption{Traditional vs Semantic Search}
\label{tab:compact_search_comparison}
\end{table}


\end{frame}

\begin{frame}
\frametitle{Resolving User Intent Ambiguity}
One primary challenge is the ambiguity of user intentions. For example, the query "Apple" could refer to different entities. Techniques like natural language processing help disambiguate these queries.
\end{frame}

\begin{frame}
\frametitle{Leveraging Knowledge Graphs}
Knowledge graphs like Google's Knowledge Graph provide structured information, enhancing the understanding of queries and contexts, thus improving search results accuracy.
\end{frame}

\begin{frame}
\frametitle{Adapting to Dynamic Language and Context}
Semantic search faces the challenge of adapting to the dynamic nature of language and evolving contexts. Constant updates and refinements are necessary for maintaining relevance.
\end{frame}

\begin{frame}
\frametitle{NLP in Semantic Search}
Natural Language Processing (NLP) enables semantic search systems to interpret the complexities of human language, facilitating entity recognition and sentiment analysis.
\begin{center} 
Process Diagram: Handling a User Query.
\end{center} 
\begin{center} 

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=2cm, minimum height=0.8cm, text centered, draw=black, fill=red!30, font=\scriptsize]
\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=0.8cm, text centered, draw=black, fill=orange!30, font=\scriptsize]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{tikzpicture}[node distance=1cm]
\node (start) [startstop] {User Query};
\node (nlp) [process, below of=start] {NLP Techniques};
\node (context) [process, below of=nlp] {Context Analysis};
\node (disambiguate) [process, below of=context] {Disambiguation Process};
\node (results) [startstop, below of=disambiguate] {Search Results};

\draw [arrow] (start) -- (nlp);
\draw [arrow] (nlp) -- (context);
\draw [arrow] (context) -- (disambiguate);
\draw [arrow] (disambiguate) -- (results);

\end{tikzpicture}
\end{center}

\end{frame}

\begin{frame}
\frametitle{Future Developments}
The future of semantic search promises more sophisticated search results, with advancements in NLP, machine learning, and knowledge graphs.
\end{frame}

\begin{frame}
\frametitle{Challenges in User Privacy and Data Security}
With the increase in user data processing, ensuring privacy and data security is paramount. This necessitates robust security measures in semantic search systems.
\end{frame}

\begin{frame}
\frametitle{Scalability and Efficiency}
Ensuring the scalability and efficiency of semantic search systems is crucial as they handle an increasing number of queries and vast amounts of data.
\end{frame}

\begin{frame}
\frametitle{Conclusion}
Semantic search, while offering improved search accuracy, faces significant challenges in understanding user intent and context. Addressing these is crucial for its future development.
\end{frame}

\end{document}
